### 2020-07-16 paper.tex
Add some more clarification to dictionaries section

Involutive MCMC is a general framework for constructing MCMC kernels that are stationary
for a target probability measure $p$. Informally, the algorithm works as follows: starting at some state $x$, we first sample
an auxiliary variable $y \sim q_x$ from a state-dependent auxiliary distribution. We then apply an \textit{involution} $f$ to the pair $(x, y)$ to obtain $(x', y')$. 
Finally, we computes an acceptance probability $\alpha$ and either accept $x'$ as the new state, or reject it and repeat the previous state $x$.
Different choices of $q$ and $f$ recover many algorithms from the literature~\citep{imcmc}.

In this section, we derive the involutive MCMC algorithm for models $p$ and auxiliary kernels
$q$ defined over general state spaces.
By emphasizing general state spaces, we intend to clarify
a potential point of confusion regarding the involutive MCMC algorithm: as presented by
\citet{imcmc},
the acceptance probability $\alpha$ depends on the Jacobian of the involution $f$, but
it is not immediately clear how to define this Jacobian when $f$ may operate on samples
from arbitrary measurable spaces, rather than on vectors in $\mathbb{R}^n$. Our 
reformulation of the algorithm below is general enough to handle arbitrary model and auxiliary
distributions, and precise enough to enable automation via probabilistic and differentiable
programming: the rest of this paper uses it to develop a technique for deriving efficient implementations
of involutive MCMC algorithms automatically, 
given only declarative specifications of $p$, $q$, and $f$.

\subsection{Distributions on Dictionaries}

Instead of vectors of real numbers, we consider in this section distributions over 
\textit{finite dictionaries} that map (possibly random) keys to (possibly random) values. 
Dictionaries include vectors as a special case (a vector $\mathbf{x} \in \mathbb{R}^n$ can
be represented as a dictionary mapping the keys $1, \dots, n$ to the values $x_1, \dots, x_n$),
but are in general more flexible: different keys can hold values of different types, and we
can also consider distributions in which the set of keys is itself random, useful for model
selection problems and structure uncertainty more generally.

\textbf{The space of finite dictionaries.}
We fix a countably infinite set $\mathcal{A}$ of possible keys, such that each key $a$ is either \emph{discrete} ($a \in \mathcal{I}$) or \emph{continuous} ($a \in \mathcal{J}$), where $\mathcal{A} = \mathcal{I} \cup \mathcal{J}$.\footnote{
It is straightforward to assign a general measure space to each address, but this is not necessary for our purposes.}
Let $V_a$ denote the set of possible values for key $a$, where $V_a$ is a countable set for each discrete key, and where $V_a = \mathbb{R}^{d_a}$ for each continuous key for some $d_a$.
Given a set of keys $A$, let $V_A = \times_{a \in A} V_a$ denote the set of \textit{assignments} of values to each key.
Then the set of all finite dictionaries is $\mathcal{D} := \bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in V_A\}$. That is, a dictionary specifies 
a finite set of keys $A \subset \mathcal{A}$ at which it has values, and an assignment $\mathbf{x}$ of values $x_a$ for each.

\textbf{A measure space of finite dictionaries.}
We associate a measure $\mu_a$ on $V_a$ for each key---the counting measure for each discrete key and the Lebesgue-measure on $\mathbf{R}^{d_a}$ for each continuous key.
For each finite set of keys $A$, we make $V_A$ a measure space using the standard product $\sigma$-algebra $\Sigma_A = \otimes_{a \in A} \Sigma_a$ and the product measure $\mu_A = \times_{a \in A} \mu_a$.
We equip $\mathcal{D}$ with the $\sigma$-algebra $\Sigma_{\mathcal{D}} := \{\bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in B_A\} \mid B_A \in \Sigma_A \, \text{for each finite } A \subset \mathcal{A}\}$ to obtain a measurable space
of dictionaries.
A reference measure $\mu_\mathcal{D}$ on this space can be constructed using the product measures $\mu_A$: we set $\mu(B) := \sum_{A \in \mathcal{A}, |A| < \infty} \mu_A(\{\mathbf{x} \mid (A, \mathbf{x}) \in B\})$.

@@

\textbf{Probability distributions on finite dictionaries.} 
When all keys are discrete, a probability distribution on dictionaries is defined by a $p$ is a probability mass function $p : \mathcal{D} \to [0, 1]$ that assigns a probability $p(m)$ to each dictionary $m \in \mathcal{P}$ such that $\sum_{x \in \mathcal{D}} p(m) = 1$.
More generally a probability distribution in dictionaries is defined by a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. We now give several examples
to build intuition.

% TODO keys vs addresses

% TODO how to notate dictionaries -- 'm' or 'd' or 'x' or 't'?

\textit{Gaussian mixture model with unknown number of components.} Consider a generative model of univariate data points $y_1, \dots, y_n$ from a Gaussian mixture with an unknown number of components $k$, each with unknown mean $m_i$ and variance $s_i$. If we place a Gaussian prior on $m_i$, an inverse Gamma prior on $s_i$, and a Poisson prior on $k$, the resulting density is
%
\begin{align*}
p(d) =
\begin{array}{l}
p_{\mathrm{poisson}(3)}(d[\mathtt{k}]) \cdot\\
\prod_{i=1}^{d[\mathtt{k}]} p_{\mathrm{normal}(0, 1)}(d[\mathtt{m}_i]) \cdot\\
\prod_{i=1}^{d[\mathtt{k}]} p_{\mathrm{inversegamma}(1, 10)}(d[\mathtt{s}_i]) \cdot\\
\prod_{i=1}^n  \frac{1}{d[\mathtt{k}]} \sum_{j=1}^{d[\mathtt{k}]} \cdot p_{\mathrm{normal}(d[\mathtt{m}_j], d[\mathtt{s}_i])}(d[\mathtt{y}_i])
\end{array}
\end{align*}
%
when $A_d$ is of the form $\{\texttt{k}, \texttt{y}_\texttt{1}, \dots, \texttt{y}_\texttt{n}\} \cup \{\texttt{m}_i \mid 1 \leq i \leq d[\texttt{k}]\} \cup \{\texttt{s}_i \mid 1 \leq i \leq d[\texttt{k}]\}$, and 0 otherwise. In this case, we have $V_\texttt{k} = \mathbb{N}$ with the counting measure for $\mu_\texttt{k}$, and for all other addresses $a \in \mathcal{A}$, $V_a = \mathbb{R}$ with the Lebesgue measure for $\mu_a$.

\textit{Polynomial regression.}

TODO

\textbf{Conditional distributions via disintegration.} Consider a probability density $p$ on the space $\mathcal{D}$ of dictionaries. 
%For a set of keys $A \subset \mathcal{A}$, let $p(A) := \int 1[A_m = A] p(m) \mu_\mathcal{D}(\text{d}m)$, the probability under $p$ of a dictionary with exactly the key set $A$. 
We say a key $a \in \mathcal{A}$ \textit{almost always appears} if $\int \mathbf{1}[a \in A_m] p(m) \mu_\mathcal{D}(\text{d}m) = 1$. Suppose $B$ is a set of keys that almost always appear for $p$, and that $b = (B, \mathbf{b})$ is a dictionary with keys $B$. Furthermore, let $(A_m, \mathbf{m}) \oplus (A_n, \mathbf{n}) := (A_m \cup A_n, \mathbf{m} \times \mathbf{n})$ be the \textit{merge} of two dictionaries $m$ and $n$ defined on disjoint key sets $A_m$ and $A_n$. Then we can define the conditional density $p(d \mid b) := \mathbf{1}[A_d \cap B = \emptyset] \frac{p(d \oplus b)}{\int_{\{m \mid A_m \cap B = \emptyset\}} p(m \oplus b) \mu_\mathcal{D}(\text{d}m)}$ when the denominator is finite. If the reference measure $\mu_a$ associated with each key $a \in B$ is discrete (e.g., the counting measure), then this definition corresponds to the ordinary notion of conditioning on an event (namely, the event that a sample from $p$ agrees with the dictionary $b$ on all addresses in $B$). When this is not the case, it corresponds to a more general measure-theoretic notion called disintegration~\citep{disintegration1997}.

\subsection{Involutive MCMC}

Let $(X, \Sigma_P, \mu_P)$ and $(Y, \Sigma_Q, \mu_Q)$ denote two general measure spaces with $\sigma$-finite $\mu_P$ and $\mu_Q$. 
Involutive MCMC (Algorithm~\ref{alg:involutive-mcmc}) implements a transition kernel that is invariant for a \textit{model distribution} given by $p : X \to [0, \infty)$, a probability density over $X$ with respect to $\mu_P$. Each iteration, the algorithm first samples auxiliary variables $y \in Y$ from an \emph{auxiliary distribution} $q_x$ based on the model's current state $x$: for each $x \in X$ such that $p(x) > 0$, $q_x : Y \to [0, \infty)$ is a probability density with respect to $\mu_Q$.
% We call $p$ the \emph{model density} and we call $q_x$ \emph{auxiliary densities}.

The resulting pair $(x, y)$ of the current model state and the newly sampled auxiliary state will be an element of the joint space $Z := \{(x, y) \in X \times Y \mid p(x) q_x(y) > 0 \}$. We can equip $Z$ with the $\sigma$-algebra 
$\Sigma := \{A \cap Z \mid A \in \Sigma_P \otimes \Sigma_Q\}$ (assuming $Z$ is a $\mu_P \times \mu_Q$-measurable set), and a reference measure $\mu(A) := (\mu_P \times \mu_Q)(A)$. 
%Let $\pi(x, y) := p(x) q_x(y)$ and $Z := \{(x, y) \in X \times Y : \pi(x, y) > 0 \}$.
%Suppose that $Z \in \Sigma_P \otimes \Sigma_Q$ ($Z$ is a $\mu_P \times \mu_Q$-measurable set).
%Then $\Sigma := \{A \in \Sigma_P \otimes \Sigma_Q : A \subseteq Z\}$ is a $\sigma$-algebra on $Z$ and $\mu : \Sigma \to [0, \infty)$ with $\mu(A) := (\mu_P \times \mu_Q)(A)$ is a $\sigma$-finite measure on measurable space $(Z, \Sigma)$.

Let $f : Z \to Z$ denote an involution ($f^{-1} = f$) such that the pushforward of $\mu$ under $f$, denoted $\mu \circ f^{-1}$, is absolutely continuous with respect to $\mu$, with Radon-Nikodym derivative $d (\mu \circ f^{-1}) / d\mu : Z \to [0, \infty)$. Involutive MCMC runs $f$ on $(x, y)$ to obtain $(x', y')$, then computes an acceptance probability $\alpha$.
With probability $\alpha$, the new state $x'$ is returned; otherwise, the previous state $x$ is repeated.
%Let $y \sim q_x(\cdot)$ denote a random sample from the measure defined by $A \mapsto \int_A q_x(y) \mu_Q(dy)$ for $A \in \Sigma_Q$.
\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary state}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot \left( \frac{d (\mu \circ f^{-1})}{d \mu} (x, y)\right)$
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC}
\label{alg:involutive-mcmc}
\end{algorithm}

@@

The general measure-theoretic formulation of involutive MCMC in Algorithm~\ref{alg:involutive-mcmc} does not specify how to compute the Radon-Nikodym derivative that is required for the acceptance probability.
When the algorithm is specialized for state spaces consisting of dictionaries, the acceptance probability simplifies and can be computed from standard Jacobian and probability density computations.
Because dictionaries are flexible objects, the algorithm remains quite general.

Suppose that the model distribution $p$ and auxiliary distributions $q_x$ are all probability distributions on dictionaries.
Then, $X$ and $Y$ are both sets of dictionaries, and the joint space $Z$ is a set of pairs $(x, y)$ of dictionaries with keys $\mathcal{A}$, so that $Z = X \times Y \subseteq \mathcal{D} \times \mathcal{D}$.
%And suppose that there exists an involution $f : Z \to Z$ such that (i) 
Suppose there is a countable partition of $Z$ into $\{Z_e : e \in E\}$ such that across each $Z_e$ the sets of keys $A_x$ and $A_y$ and the values $x_a$ and $y_a$ for all discrete keys $a$ are both constant (so that each set $Z_e$ is isomorphic to a Euclidean space of assignments to the continuous keys in the two dictionaries).
Suppose there is an involution $g : E \to E$ between elements of the partition (a function with $g = g^{-1}$), and a family of continuously differentiable bijections $h_{e} : Z_e \to Z_{g(e)}$ for each $e$ with $h_{e} = h_{g(e)}^{-1}$.
Let $e(z) \in E$ denote the element of the partition for an alement $z \in Z$.
Then, $f : Z \to Z$ given by $f(z) := h_{e(z)}(z)$ is an involution:
\begin{equation}
f(f(z)) = h_{e(f(z))}(h_{e(z)}(z)) = h_{g(e(z))}(h_{e(z)}(z)) = z
\end{equation}
Let $|J h_e|(z)$ denote the absolute value of the Jacobian (determinant) of $h_e$, evaluated at $z$.
Then, the acceptance probability simplifies to:
\begin{equation}
\alpha(x, y) = \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot |J h_e|(z)
\end{equation}
Section~\ref{sec:radon-nikodym-special-case} of the appendix for details.

>>>>

Involutive MCMC is a general framework for constructing MCMC kernels that are stationary
for a target probability distribution $p$. Informally, the algorithm works as follows: starting at some state $x$, we first sample
an auxiliary variable $y \sim q_x$ from a state-dependent auxiliary distribution. We then apply an \textit{involution} $f$ to the pair $(x, y)$ to obtain $(x', y')$. 
Finally, we computes an acceptance probability $\alpha$ and either accept $x'$ as the new state, or reject it and repeat the previous state $x$.
Different choices of $q$ and $f$ recover many algorithms from the literature~\citep{imcmc}.

In this section, we present involutive MCMC for models $p$ and auxiliary kernels
$q$ defined over general state spaces.
By emphasizing general state spaces, we intend to clarify
a potential point of confusion regarding the involutive MCMC algorithm: as presented by
\citet{imcmc},
the acceptance probability $\alpha$ depends on the Jacobian of the involution $f$, but
it is not immediately clear how to define this Jacobian when $f$ may operate on samples
from arbitrary measurable spaces, rather than on vectors in $\mathbb{R}^n$. Our 
reformulation of the algorithm below is general enough to handle arbitrary model and auxiliary
distributions, and precise enough to enable automation via probabilistic and differentiable
programming: the rest of this paper uses it to develop a technique for deriving efficient implementations
of involutive MCMC algorithms automatically, 
given only declarative specifications of $p$, $q$, and $f$.

\subsection{General Involutive MCMC}

Let $(X, \Sigma_P, \mu_P)$ and $(Y, \Sigma_Q, \mu_Q)$ denote two general measure spaces with $\sigma$-finite $\mu_P$ and $\mu_Q$. 
Involutive MCMC (Algorithm~\ref{alg:involutive-mcmc}) implements a transition kernel that is invariant for a \textit{model distribution} given by $p : X \to [0, \infty)$, a probability density over $X$ with respect to $\mu_P$. Each iteration, the algorithm first samples auxiliary variables $y \in Y$ from an \emph{auxiliary distribution} $q_x$ based on the model's current state $x$: for each $x \in X$ such that $p(x) > 0$, $q_x : Y \to [0, \infty)$ is a probability density with respect to $\mu_Q$.

The resulting pair $(x, y)$ of the current model state and the newly sampled auxiliary state will be an element of the joint space $Z := \{(x, y) \in X \times Y \mid p(x) q_x(y) > 0 \}$. We can equip $Z$ with the $\sigma$-algebra 
$\Sigma := \{A \cap Z \mid A \in \Sigma_P \otimes \Sigma_Q\}$ (assuming $Z$ is a $\mu_P \times \mu_Q$-measurable set), and a reference measure $\mu(A) := (\mu_P \times \mu_Q)(A)$.

Let $f : Z \to Z$ denote an involution ($f^{-1} = f$) such that the pushforward of $\mu$ under $f$, denoted $\mu \circ f^{-1}$, is absolutely continuous with respect to $\mu$, with Radon-Nikodym derivative $d (\mu \circ f^{-1}) / d\mu : Z \to [0, \infty)$. Involutive MCMC runs $f$ on $(x, y)$ to obtain $(x', y')$, then computes an acceptance probability $\alpha$.
With probability $\alpha$, the new state $x'$ is returned; otherwise, the previous state $x$ is repeated.

\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary state}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot \left( \frac{d (\mu \circ f^{-1})}{d \mu} (x, y)\right)$
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC}
\label{alg:involutive-mcmc}
\end{algorithm}

\subsection{Probability Distributions on Dictionaries}

While maximally general, the measure-theoretic formulation of involutive MCMC in Algorithm~\ref{alg:involutive-mcmc} is not amenable to an automated implementation, because it does not indicate how to compute the Radon-Nikodym derivative that is required for the acceptance probability, and it is unclear how to specify the probability measures involved.
%When the algorithm is specialized for state spaces consisting of dictionaries, the acceptance probability simplifies and can be computed from standard Jacobian and probability density computations.
%Because dictionaries are flexible objects, the algorithm remains quite general.

While restricting the state space to vectors of real numbers would address these issues, we seek a 
representation that remains flexible enough to represent complex hybrid state spaces with numeric, symbolic, and structure uncertainty.
Therefore, we use state spaces consisting of 
\textit{finite dictionaries} that map (possibly random) keys to (possibly random) values. 
Dictionaries include vectors as a special case (a vector $\mathbf{x} \in \mathbb{R}^n$ can
be represented as a dictionary mapping the keys $1, \dots, n$ to the values $x_1, \dots, x_n$),
but are more flexible: different keys can hold values of different types (e.g. integers, strings), and we
can also consider distributions in which the set of keys is itself random, which is useful for model
selection problems and structure uncertainty more generally.

This section describes probability distributions on dictionaries, and gives a constructive definition of the involutive MCMC acceptance probability in this setting in terms of a Jacobian.
Section~\ref{sec:automating} will then show how probability distributions on dictionaries can be specified with probabilistic programs, and how probabilistic programming techniques can automatically compute probability densities spaces of dictionaries.

\textbf{The space of finite dictionaries.}
We fix a countably infinite set $\mathcal{A}$ of possible keys, such that each key $a$ is either called \emph{discrete} ($a \in \mathcal{I}$) or \emph{continuous} ($a \in \mathcal{J}$), where $\mathcal{A} = \mathcal{I} \cup \mathcal{J}$.\footnote{
It is possible to assign a general measure space to each address, but this is not necessary for our purposes.}
Let $V_a$ denote the set of possible values for key $a$, where $V_a$ is a countable set for each discrete key, and where $V_a = \mathbb{R}^{d_a}$ for each continuous key for some $d_a$.
Given a set of keys $A$, let $V_A = \times_{a \in A} V_a$ denote the set of \textit{assignments} of values to each key.
Then the set of all finite dictionaries is $\mathcal{D} := \bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in V_A\}$. That is, a dictionary specifies 
a finite set of keys $A \subset \mathcal{A}$ at which it has values, and an assignment $\mathbf{x}$ of values $x_a$ for each.

\textbf{Relationship to representation of~\citet{green1995reversible}.}
\citet{green1995reversible} uses a state space that is the countable union of `models', where for each model there is a Eucliean space of parameters.
Dictionaries possesses substantially more structure:
Parameters have keys, so it is possible to express that parameters are shared across different structures.
Instead of a monolithic `model', dictionaries use a more elaborate discrete state that includes the set of keys and the assignment to the discrete keys.
The additional structure of dictionaries enables the automation technique Section~\ref{sec:automating}.

\textbf{A measure space of finite dictionaries.}
We associate a measure $\mu_a$ on $V_a$ for each key---the counting measure for each discrete key and the Lebesgue-measure on $\mathbb{R}^{d_a}$ for each continuous key.
For each finite set of keys $A$, we make $V_A$ a measure space using the standard product $\sigma$-algebra $\Sigma_A = \otimes_{a \in A} \Sigma_a$ and the product measure $\mu_A = \times_{a \in A} \mu_a$.
We equip $\mathcal{D}$ with the $\sigma$-algebra $\Sigma_{\mathcal{D}} := \{\bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in B_A\} \mid B_A \in \Sigma_A \, \text{for each finite } A \subset \mathcal{A}\}$ to obtain a measurable space
of dictionaries.
A reference measure $\mu_\mathcal{D}$ on this space can be constructed using the product measures $\mu_A$: we set $\mu(B) := \sum_{A \in \mathcal{A}, |A| < \infty} \mu_A(\{\mathbf{x} \mid (A, \mathbf{x}) \in B\})$.

@@

\textbf{Probability distributions on finite dictionaries.} 
When all keys are discrete, a probability distribution on dictionaries is defined by a $p$ is a probability mass function $p : \mathcal{D} \to [0, 1]$ that assigns a probability $p(m)$ to each dictionary $m \in \mathcal{P}$ such that $\sum_{x \in \mathcal{D}} p(m) = 1$.
More generally a probability distribution on dictionaries is defined by a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. 
The probability is distributed the finite sets of keys $A \subseteq \mathcal{A}$ where $|A| < \infty$:
\begin{equation}
1 = \int p(m) \mu_\mathcal{D}(\text{d}m) = \sum_{\substack{A \subseteq \mathcal{A}\\|A| < \infty}} \int p(m) \mu_{A}(dm)
\end{equation}
where the probability for each key set $A$ is:
\begin{equation}\label{eq:integral}
\int p(m) \mu_{A}(dm) = \sum_{\mathbf{x}_1} \int_{\mathbb{R}^{d}} p((A, (\mathbf{x}_1, \mathbf{x}_2))) d\mathbf{x}_2
\end{equation}
where $d: = \sum_{a \in A \cap \mathcal{J}} d_{a}$ and where $\mathbf{x}_1$ is an assignment to the discrete choices in $A$ and $\mathbf{x}_2$ is an assignment to the continuous choices.

We now give an example to build intuition.
Consider a generative model of univariate data points $y_1, \dots, y_n$ from a Gaussian mixture with an unknown number of components $k$, each with unknown mean $m_i$ and variance $s_i$. If we place a Gaussian prior on $m_i$, an inverse Gamma prior on $s_i$, and a Poisson prior on $k$, the resulting density on dictionaries $d$ is:
%
\begin{align*}
p(d) =
\begin{array}{l}
p_{\mathrm{poisson}(3)}(d[\mathtt{k}]) \cdot\\
\prod_{i=1}^{d[\mathtt{k}]} p_{\mathrm{normal}(0, 1)}(d[\mathtt{m}_i]) \cdot\\
\prod_{i=1}^{d[\mathtt{k}]} p_{\mathrm{inversegamma}(1, 10)}(d[\mathtt{s}_i]) \cdot\\
\prod_{i=1}^n  \frac{1}{d[\mathtt{k}]} \sum_{j=1}^{d[\mathtt{k}]} \cdot p_{\mathrm{normal}(d[\mathtt{m}_j], d[\mathtt{s}_i])}(d[\mathtt{y}_i])
\end{array}
\end{align*}
%
when $A_d = \{\texttt{k}, \texttt{y}_\texttt{1}, \dots, \texttt{y}_\texttt{n}\} \cup \{\texttt{m}_i \mid 1 \leq i \leq d[\texttt{k}]\} \cup \{\texttt{s}_i \mid 1 \leq i \leq d[\texttt{k}]\}$, and 0 otherwise. In this case, we have $V_\texttt{k} = \mathbb{N}$ with the counting measure for $\mu_\texttt{k}$, and for all other addresses $a \in \mathcal{A}$, $V_a = \mathbb{R}$ with the Lebesgue measure for $\mu_a$.
For each $k \in \{0, 1, \ldots\}$, the probability for key set
$\{\texttt{k}, \texttt{y}_\texttt{1}, \dots, \texttt{y}_\texttt{n}\} \cup \{\texttt{m}_i \mid 1 \leq i \leq k\} \cup \{\texttt{s}_i \mid 1 \leq i \leq k\}$
is $p_{\mathrm{poisson}(s)}(k)$ (via Equation~(\ref{eq:integral})).

\textbf{Conditional distributions via disintegration.} Consider a probability density $p$ on the space $\mathcal{D}$ of dictionaries. 
%For a set of keys $A \subset \mathcal{A}$, let $p(A) := \int 1[A_m = A] p(m) \mu_\mathcal{D}(\text{d}m)$, the probability under $p$ of a dictionary with exactly the key set $A$. 
We say a key $a \in \mathcal{A}$ \textit{almost always appears} if $\int \mathbf{1}[a \in A_m] p(m) \mu_\mathcal{D}(\text{d}m) = 1$. Suppose $B$ is a set of keys that almost always appear for $p$, and that $b = (B, \mathbf{b})$ is a dictionary with keys $B$. Furthermore, let $(A_m, \mathbf{m}) \oplus (A_n, \mathbf{n}) := (A_m \cup A_n, \mathbf{m} \times \mathbf{n})$ be the \textit{merge} of two dictionaries $m$ and $n$ defined on disjoint key sets $A_m$ and $A_n$. Then we can define the conditional density $p(d \mid b) := \mathbf{1}[A_d \cap B = \emptyset] \frac{p(d \oplus b)}{\int_{\{m \mid A_m \cap B = \emptyset\}} p(m \oplus b) \mu_\mathcal{D}(\text{d}m)}$ when the denominator is finite. If the reference measure $\mu_a$ associated with each key $a \in B$ is discrete (e.g., the counting measure), then this definition corresponds to the ordinary notion of conditioning on an event (namely, the event that a sample from $p$ agrees with the dictionary $b$ on all addresses in $B$). When this is not the case, it corresponds to a more general measure-theoretic notion called disintegration~\citep{chang1997conditioning}.

Consider the infinite univariate mixture model, and the conditional density given observed data $(B, \mathbf{b}) := \{\mathtt{y}_1 \mapsto y_1, \ldots, \mathtt{y}_{\mathtt{n}} \mapsto y_{\mathtt{n}}\}$.
The conditional density $p(d | b)$ is nonzero only if $A_d = \{\texttt{k}\} \cup \{\texttt{m}_i \mid 1 \leq i \leq k\} \cup \{\texttt{s}_i \mid 1 \leq i \leq k\}$ for some $k$ ($d$ does not contain y-values), and the denominator in the definition of $p(d | b)$ simplifies to the familiar sum of marginal likelihoods over all $k$, where each marginal likelihood is an integral over $\mathbb{R}^{2k}$.

@@

%The general measure-theoretic formulation of involutive MCMC in Algorithm~\ref{alg:involutive-mcmc} does not specify how to compute the Radon-Nikodym derivative that is required for the acceptance probability.
%When the algorithm is specialized for state spaces consisting of dictionaries, the acceptance probability simplifies and can be computed from standard Jacobian and probability density computations.
%Because dictionaries are flexible objects, the algorithm remains quite general.

Suppose that the model distribution and auxiliary distributions are probability distributions on dictionaries, with densities $p$ and $q_x$.
Then, $X$ and $Y$ are both sets of dictionaries, and the joint space $Z$ is a set of pairs $(x, y)$ of dictionaries with keys $\mathcal{A}$, so that $Z = X \times Y \subseteq \mathcal{D} \times \mathcal{D}$.
%And suppose that there exists an involution $f : Z \to Z$ such that (i) 
Suppose there is a countable partition of $Z$ into $\{Z_e : e \in E\}$ such that across each $Z_e$ the sets of keys $A_x$ and $A_y$ and the values $x_a$ and $y_a$ for all discrete keys $a$ are both constant (so that each set $Z_e$ is isomorphic to a Euclidean space of assignments to the continuous keys in the two dictionaries).
Suppose there is an involution $g : E \to E$ between elements of the partition (a function with $g = g^{-1}$), and a family of continuously differentiable bijections $h_{e} : Z_e \to Z_{g(e)}$ for each $e$ with $h_{e} = h_{g(e)}^{-1}$.
Let $e(z) \in E$ denote the element of the partition for an alement $z \in Z$.
Then, $f : Z \to Z$ given by $f(z) := h_{e(z)}(z)$ is an involution:
\begin{equation}
f(f(z)) = h_{e(f(z))}(h_{e(z)}(z)) = h_{g(e(z))}(h_{e(z)}(z)) = z
\end{equation}
Let $|J h_e|(z)$ denote the absolute value of the Jacobian (determinant) of $h_e$, evaluated at $z$.
Then, the acceptance probability simplifies to:
\begin{equation}
\alpha(x, y) = \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot |J h_e|(z)
\end{equation}
Section~\ref{sec:radon-nikodym-special-case} of the appendix for details.
